---
title: "assigment Session 7"
author:
  name: Jan Pamula
format: 
  html:
      embed-resources: true
      theme: lumen
      toc: true
      toc-color: red
      toc-location: left
      toc-depth: 6
      toc-expand: 6
      code-tools:
        source: true
        toggle: true
      code-fold: true
      code-summary: "Show the code"
      code-overflow: wrap
      page-layout: full
editor: visual
editor_options: 
  markdown: 
    wrap: 72
---

# Loading libraries

```{r}
library(ggplot2)
```

# First plot

creating x and y axis variables and data-frame:

```{r}
first_x <- c(2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024, 2025, 2026, 2027)
first_y <- c(0.048, 0.047, 0.045, 0.044, 0.046, 0.044, 0.044, 0.045, 0.0463, 0.0475, 0.0488, 0.05, 0.053, 0.0565, 0.06, 0.0625, 0.065, 0.0675, 0.07)
first_data <- data.frame(
  year = first_x,
  spend = first_y
)
```

```{r}
p<-ggplot(data=first_data, aes(x=factor(year), y=spend)) +
  geom_bar(stat="identity",  aes(fill = year > 2023)) +
  geom_text(aes(label=scales::percent(spend)), vjust=-0.3, size=2) +
  scale_fill_discrete(name = "", labels=c('wydatki', "przewdziane \nprzyszłe wydatki")) +
  ggtitle("Wydatki na ochorone zdrowia jakko % PKB")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90, size = 10))
p
```

-   `p<-` - assigning plot to variable (called "p")
-   `ggplot(data=first_data, aes(x=factor(year), y=spend))` - assign
    data to ggplot (x-axis = first_data.year, y-axis = first_data.spend)
-   `geom_bar(stat="identity",  aes(fill = year > 2023))` - creating bar
    chart where bars(years) \> 2023 are colored for different color
-   `geom_text(aes(label=scales::percent(spend)), vjust=-0.3, size=2)` -
    displaying every value on bars (in percentage)
-   ggtitle("Wydatki na ochorone zdrowia jakko % PKB") - changing title
-   `scale_fill_discrete(name = "", labels=c('wydatki', "przewdziane \nprzyszłe wydatki"))` -
    changing legend description
-   `theme_minimal()` - using minimal theme
-   `theme(plot.title = element_text(hjust = 0.5),` - centering title
-   `axis.text.x = element_text(angle = 90, size = 10))` - rotating
    x-axis values

# Second plot

creating x and y axis variables and data-frame:

```{r}
second_x <- c("China", "US", "EEA", "India", "Russia", "Japan", "Other")
second_y <- c(0.294, 0.143, 0.098, 0.068, 0.049, 0.035, 0.315)
second_data <- data.frame(
  Country = second_x,
  Emission = second_y
)
```

```{r}
q<-ggplot(data=second_data, aes(x=Country, y=Emission, fill = Country)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(Emission)), vjust=-0.3, size=3.5) +
  ggtitle("CO2 Emsissions")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
q
```

-   `q<-` - assigning plot to variable (called "p")
-   `ggplot(data=second_data, aes(x=Country, y=Emission, fill = Country))` -
    assign data to ggplot (x-axis = second_data.Country, y-axis =
    second_data.Emission) and making different color for every country
    (`fill = Country`)
-   `geom_bar(stat="identity")` - creating bar chart
-   `geom_text(aes(label=scales::percent(Emission)), vjust=-0.3, size=3.5)` -
    displaying every value on bars (in percentage)
-   `ggtitle("CO2 Emsissions")` - changing title
-   `theme_minimal()` - using minimal theme
-   `theme(plot.title = element_text(hjust = 0.5))` - centering title
